---
title: "HW 11"
author: "Name: Paul Holaway  , NetID: paulch2"
date: 'Due: 11/10/2022 11:59pm'
output: pdf_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(astsa)
library(latex2exp)
```
### Question 1.

Find A and B so that 
$$3 \cos (2\pi \omega t+0.4) = A\cos (2\pi \omega t)+ B\sin (2\pi \omega t).$$

$$
3\cos(2\pi\omega t+0.4)=3\cos(2\pi\omega t)\cos(0.4)-3\sin(2\pi\omega t)\sin(0.4)
$$
$$
=3\cos(0.4)\cos(2\pi\omega t)-2\sin(0.4)\sin(2\pi\omega t)
$$
$$
\implies\boxed{A=3\cos(0.4)\text{, }B=-3\sin(0.4)}
$$

### Question 2. 

Find R and $\Phi$ so that 
$$ R \cos (2\pi \omega t+\Phi)= \cos (2\pi \omega t)+ 3\sin (2\pi \omega t).$$
$$
=R\cos(2\pi\omega t)cos(\Phi)-R\sin(2\pi\omega t)\sin(\Phi)
$$
$$
R\cos(\Phi)=1\text{ and }-R\sin(\Phi)=3
$$
$$
\implies R=\frac{1}{\cos(\Phi)}\implies-\frac{\sin(\Phi)}{\cos(\Phi)}=3\implies-\tan(\Phi)=3\implies\Phi=\tan^{-1}(-3)\implies R=\sec(\tan^{-1}(-3))
$$
$$
\boxed{\Phi=\tan^{-1}(-3)\text{, }R=\sec(\tan^{-1}(-3))}
$$

\newpage

### Question 3. 

Let $$X_t=2 \cos \left(2\pi t \frac{4}{96} \right)+2 \sin \left(2\pi  \left(t\frac{14}{96}+0.3\right) \right).$$

(a) Make a time series plot of the time series, for $t=1, \dots, 96$.

```{r}
X1 = 2*cos(2*pi*1:96*(4/96)) + 2*sin(2*pi*(1:96*(14/96)+0.3))
tsplot(X1, main="")
```

\newpage

(b) Conduct the regression of $X_t$ on $\cos (2\pi \omega t)$ and $\sin(2\pi \omega t)$ for $\omega=\frac{4}{96}$. Use R. Verify that they are perfect estimates (no error/noise term).

```{r}
#w = 4/96
Z1 = cos(2*pi*(4/96)*1:96)
Z2 = sin(2*pi*(4/96)*1:96)
#Regression
m1 = lm(X1 ~ Z1 + Z2)
#Model
summary(m1)
```
Yes, we have perfect estimates. We notice that the $R^2$ is exactly 0.5, our `Z1` coefficient is 2, and our `Z2` coefficient is 0. This means that we are perfectly explaining the first part of $X_t$ when $\omega=\frac{4}{96}$.

\newpage

(c) Conduct the regression of $X_t$ on $\cos (2\pi \omega t)$ and $\sin(2\pi \omega t)$ for $\omega=\frac{14}{96}$. Use R. Verify that they are perfect estimates (no error/noise term).

```{r}
#w = 14/96
Z3 = cos(2*pi*(14/96)*1:96)
Z4 = sin(2*pi*(14/96)*1:96)
#Regression
m2 = lm(X1 ~ Z3 + Z4)
#Model
summary(m2)
```
Yes, we have perfect estimates. We notice that the $R^2$ is exactly 0.5, our `Z3` coefficient is 1.902, and our `Z4` coefficient is -0.618. This means that we are perfectly explaining the second part of $X_t$ when $\omega=\frac{14}{96}$.

\newpage

(d) Conduct the regression of $X_t$ on $\cos (2\pi \omega t)$ and $\sin(2\pi \omega t)$ for $\omega=\frac{4}{96}$ and $\omega=\frac{14}{96}$ together. Use R. Verify that they are perfect estimates (no error/noise term).

```{r}
#w = 4/96 and 14/96 together
#Regression
m3 = lm(X1 ~ Z1 + Z2 + Z3 + Z4)
#Model
summary(m3)
```
Yes, we have perfect estimates. We notice that the $R^2$ is exactly 1, which means that we are perfectly explaining both parts of $X_t$.

\newpage

(e) Conduct the regression of $X_t$ on $\cos (2\pi \omega t)$ and $\sin(2\pi \omega t)$ for $\omega=\frac{3}{96}$ and $\omega=\frac{13}{96}$ together. Use R. Are those estimates still perfect?

```{r}
#w = 3/96 and 13/96 together
Z5 = cos(2*pi*(3/96)*1:96)
Z6 = sin(2*pi*(3/96)*1:96)
Z7 = cos(2*pi*(13/96)*1:96)
Z8 = sin(2*pi*(13/96)*1:96)
#Regression
m4 = lm(X1 ~ Z5 + Z6 + Z7 + Z8)
#Model
summary(m4)
```
No, the estimates are no longer perfect. We can see that the $R^2$ is practically 0 and all of the coefficient estimates are also practically 0. Note also how none of them are insignificant.

\newpage

### Question 4. 

Generate any series of length n = 10. Show that the series may be fit exactly by a linear combination of enough cosine-sine curves at the Fourier frequencies, $\omega=1/10, \dots, 5/10$. You may use R and conduct regression for this problem. 

```{r}
#Creating Sequence
myseq = rep(0,10)

for(i in 1:10){
  myseq[i] = 2^i
}

#Creating Frequency Terms
F1 = cos(2*pi*(1/10)*1:10)
F2 = sin(2*pi*(1/10)*1:10)
F3 = cos(2*pi*(2/10)*1:10)
F4 = sin(2*pi*(2/10)*1:10)
F5 = cos(2*pi*(3/10)*1:10)
F6 = sin(2*pi*(3/10)*1:10)
F7 = cos(2*pi*(4/10)*1:10)
F8 = sin(2*pi*(4/10)*1:10)
F9 = cos(2*pi*(5/10)*1:10)
F10 = sin(2*pi*(5/10)*1:10)

#Creating models
M1 = lm(myseq ~ F1 + F2)
M2 = lm(myseq ~ F1 + F2 + F3 + F4)
M3 = lm(myseq ~ F1 + F2 + F3 + F4 + F5 + F6)
M4 = lm(myseq ~ F1 + F2 + F3 + F4 + F5 + F6 + F7 + F8)
M5 = lm(myseq ~ F1 + F2 + F3 + F4 + F5 + F6 + F7 + F8 + F9 + F10)

#Plotting
plot(1:10, myseq, col = "white", xlab = "", ylab = "Series", ylim = c(-100,1000))
grid()
box()
lines(1:10, myseq)
lines(1:10, M1$fitted.values, col = "red")
lines(1:10, M2$fitted.values, col = "yellow")
lines(1:10, M3$fitted.values, col = "green")
lines(1:10, M4$fitted.values, col = "blue")
lines(1:10, M5$fitted.values, col = "violet")
legend("topleft", legend = c("Series",TeX(r"($\omega$=1/10)"),TeX(r"($\omega$=1/10:2/10)"),
       TeX(r"($\omega$=1/10:3/10)"), TeX(r"($\omega$=1/10:4/10)"),TeX(r"($\omega$=1/10:5/10)")),
       col = c("black","red","yellow","green","blue","violet"), lty = 1)
```
As we can see, once we use all five of the Fourier frequencies ($\omega=1/10, \ldots, 5/10$), the series of $2^x$ will be fit exactly by a linear combination of the cos-sin curves.

\newpage

### Question 5. [GR-only]

For simplicity, let's assume that $n$ is an even integer.

Hint: Use $\cos (b)=\frac{e^{ib}+e^{-ib}}{2}$, $\sin (b)=\frac{e^{ib}-e^{-ib}}{2i}$, $e^{ib}=\cos(b)+ i\sin (b)$.

Verify that for any positive integer $n$ and $j,k=0,1,\dots n/2$:

(a) Except for $j=0$ or $j=n/2$, 

$$\sum_{t=1}^{n}\cos^2 (2\pi t \frac{j}{n})= \sum_{t=1}^{n}\sin^2 (2\pi t \frac{j}{n})=\frac{n}{2}.$$
$$
\sum_{t=1}^{n}\cos^2 (2\pi t \frac{j}{n})=\sum_{t=1}^n(\frac{e^{2\pi t\frac{j}{n}i}+e^{-2\pi t\frac{j}{n}i}}{2})^2=\sum_{t=1}^n(\frac{e^{4\pi t\frac{j}{n}i}+e^{-4\pi t\frac{j}{n}i}+2}{4})
$$
$$
=\sum_{t=1}^{n}[\frac{1}{2}+\frac{\cos(4\pi t\frac{j}{n})}{4}+\frac{i\sin(4\pi t\frac{j}{n})}{4}+\frac{\cos(-4\pi t\frac{j}{n})}{4}+\frac{i\sin(-4\pi t\frac{j}{n})}{4}]
$$
$$
=\sum_{t=1}^{n}[\frac{1}{2}+\frac{\cos(4\pi t\frac{j}{n})}{4}+\frac{i\sin(4\pi t\frac{j}{n})}{4}+\frac{\cos(4\pi t\frac{j}{n})}{4}-\frac{i\sin(4\pi t\frac{j}{n})}{4}]
$$
$$
=\frac{n}{2}+\frac{1}{2}\sum_{t=1}^{n}\cos(4\pi t\frac{j}{n})
$$
$$
\sum_{t=1}^{n}\cos(4\pi t\frac{j}{n})=0\text{ as }n\text{ is even.}
$$
$$
\implies \sum_{t=1}^{n}\cos^2 (2\pi t \frac{j}{n})=\frac{n}{2}
$$
$$
\sum_{t=1}^{n}\sin^2 (2\pi t \frac{j}{n})=\sum_{t=1}^n(\frac{e^{2\pi t\frac{j}{n}i}-e^{-2\pi t\frac{j}{n}i}}{2i})^2=\sum_{t=1}^n(\frac{e^{4\pi t\frac{j}{n}i}+e^{-4\pi t\frac{j}{n}i}-2}{-4})
$$
$$
=\sum_{t=1}^{n}[\frac{1}{2}-\frac{\cos(4\pi t\frac{j}{n})}{4}-\frac{i\sin(4\pi t\frac{j}{n})}{4}-\frac{\cos(-4\pi t\frac{j}{n})}{4}-\frac{i\sin(-4\pi t\frac{j}{n})}{4}]
$$
$$
=\sum_{t=1}^{n}[\frac{1}{2}-\frac{\cos(4\pi t\frac{j}{n})}{4}-\frac{i\sin(4\pi t\frac{j}{n})}{4}-\frac{\cos(4\pi t\frac{j}{n})}{4}+\frac{i\sin(4\pi t\frac{j}{n})}{4}]
$$
$$
=\frac{n}{2}-\frac{1}{2}\sum_{t=1}^{n}\cos(4\pi t\frac{j}{n})
$$
$$
\sum_{t=1}^{n}\cos(4\pi t\frac{j}{n})=0\text{ as }n\text{ is even.}
$$
$$
\implies \sum_{t=1}^{n}\sin^2 (2\pi t \frac{j}{n})=\frac{n}{2}
$$

(b) When $j=0$ or $j=n/2$, 

$$\sum_{t=1}^{n}\cos^2 (2\pi t \frac{j}{n})=n, \quad \sum_{t=1}^{n}\sin^2 (2\pi t \frac{j}{n})=0.$$

* $j=0$

$$
\sum_{t=1}^{n}\cos^2 (2\pi t \frac{j}{n})=\sum_{t=1}^{n}\cos^2 (0)=\sum_{t=1}^{n}1=n
$$
$$
\sum_{t=1}^{n}\sin^2 (2\pi t \frac{j}{n})=\sum_{t=1}^{n}\sin^2 (0)=\sum_{t=1}^{n}0=0
$$

* $j=\frac{n}{2}$

$$
\sum_{t=1}^{n}\cos^2 (2\pi t \frac{j}{n})=\sum_{t=1}^{n}\cos^2 (\pi t)=\sum_{t=1}^{n}1=n
$$
$$
\sum_{t=1}^{n}\sin^2 (2\pi t \frac{j}{n})=\sum_{t=1}^{n}\sin^2 (\pi t)=\sum_{t=1}^{n}0=0
$$

(c) When $j\neq k$, 

$$\sum_{t=1}^{n}\cos (2\pi t \frac{j}{n}) \cos (2\pi t \frac{k}{n}) =\sum_{t=1}^{n}\sin (2\pi t \frac{j}{n}) \sin (2\pi t \frac{k}{n})  =0 .$$
$$
\sum_{t=1}^{n}\cos (2\pi t \frac{j}{n}) \cos (2\pi t \frac{k}{n}) =\sum_{t=1}^{n}(\frac{e^{2\pi t\frac{j}{n}i}+e^{-2\pi t\frac{j}{n}i}}{2})(\frac{e^{2\pi t\frac{k}{n}i}+e^{-2\pi t\frac{k}{n}i}}{2})
$$
$$
=\frac{1}{4}\sum_{t=1}^{n}e^{\frac{2\pi t}{n}(j+k)i}+e^{\frac{2\pi t}{n}(k-j)i}+e^{\frac{2\pi t}{n}(j-k)i}+e^{\frac{2\pi t}{n}(-j-k)i}
$$
$$
=\frac{1}{4}\sum_{t=1}^{n}[\cos(\frac{2\pi t}{n}(j+k))+i\sin(\frac{2\pi t}{n}(j+k))+\cos(\frac{2\pi t}{n}(k-j))+i\sin(\frac{2\pi t}{n}(k-j))
$$
$$
+\cos(\frac{2\pi t}{n}(j-k))+i\sin(\frac{2\pi t}{n}(j-k))+\cos(\frac{2\pi t}{n}(-j-k))+i\sin(\frac{2\pi t}{n}(-j-k))]
$$
$$
=\frac{1}{4}\sum_{t=1}^{n}[\cos(\frac{2\pi t}{n}(j+k))+i\sin(\frac{2\pi t}{n}(j+k))+\cos(\frac{2\pi t}{n}(j-k))-i\sin(\frac{2\pi t}{n}(j-k))
$$
$$
+\cos(\frac{2\pi t}{n}(j-k))+i\sin(\frac{2\pi t}{n}(j-k))+\cos(\frac{2\pi t}{n}(j+k))-i\sin(\frac{2\pi t}{n}(j+k))]
$$
$$
=\frac{1}{2}\sum_{t=1}^{n}\cos(\frac{2\pi t}{n}(j+k))+\cos(\frac{2\pi t}{n}(j-k))
$$
$$
\sum_{t=1}^{n}\cos(\frac{2\pi t}{n}(j+k))+\cos(\frac{2\pi t}{n}(j-k))=0\text{ as }n\text{ is even.}
$$
$$
\sum_{t=1}^{n}\cos (2\pi t \frac{j}{n}) \cos (2\pi t \frac{k}{n}) =0
$$
$$
\sum_{t=1}^{n}\sin (2\pi t \frac{j}{n}) \sin (2\pi t \frac{k}{n})=\sum_{t=1}^{n}(\frac{e^{2\pi t\frac{j}{n}i}-e^{-2\pi t\frac{j}{n}i}}{2i})(\frac{e^{2\pi t\frac{k}{n}i}-e^{-2\pi t\frac{k}{n}i}}{2i})
$$
$$
=-\frac{1}{4}\sum_{t=1}^{n}e^{\frac{2\pi t}{n}(j+k)i}-e^{\frac{2\pi t}{n}(k-j)i}-e^{\frac{2\pi t}{n}(j-k)i}+e^{\frac{2\pi t}{n}(-j-k)i}
$$
$$
=-\frac{1}{4}\sum_{t=1}^{n}[\cos(\frac{2\pi t}{n}(j+k))+i\sin(\frac{2\pi t}{n}(j+k))-\cos(\frac{2\pi t}{n}(k-j))-i\sin(\frac{2\pi t}{n}(k-j))
$$
$$
-\cos(\frac{2\pi t}{n}(j-k))-i\sin(\frac{2\pi t}{n}(j-k))+\cos(\frac{2\pi t}{n}(-j-k))+i\sin(\frac{2\pi t}{n}(-j-k))]
$$
$$
=-\frac{1}{4}\sum_{t=1}^{n}[\cos(\frac{2\pi t}{n}(j+k))+i\sin(\frac{2\pi t}{n}(j+k))-\cos(\frac{2\pi t}{n}(j-k))+i\sin(\frac{2\pi t}{n}(j-k))
$$
$$
-\cos(\frac{2\pi t}{n}(j-k))-i\sin(\frac{2\pi t}{n}(j-k))+\cos(\frac{2\pi t}{n}(j+k))-i\sin(\frac{2\pi t}{n}(j+k))]
$$
$$
=\frac{1}{2}\sum_{t=1}^{n}\cos(\frac{2\pi t}{n}(j-k))-\cos(\frac{2\pi t}{n}(j+k))
$$
$$
\sum_{t=1}^{n}\cos(\frac{2\pi t}{n}(j-k))-\cos(\frac{2\pi t}{n}(j+k))=0\text{ as }n\text{ is even.}
$$
$$
\sum_{t=1}^{n}\sin (2\pi t \frac{j}{n}) \sin (2\pi t \frac{k}{n}) =0
$$

(d) For any $j$ and $k$,

$$
\sum_{t=1}^{n}\cos (2\pi t \frac{j}{n}) \sin (2\pi t \frac{k}{n})=0 .
$$
$$
\sum_{t=1}^{n}\cos (2\pi t \frac{j}{n}) \sin (2\pi t \frac{k}{n})=\sum_{t=1}^{n}(\frac{e^{2\pi t\frac{j}{n}i}+e^{-2\pi t\frac{j}{n}i}}{2})(\frac{e^{2\pi t\frac{k}{n}i}-e^{-2\pi t\frac{k}{n}i}}{2i})
$$
$$
=\frac{1}{4i}\sum_{t=1}^{n}e^{\frac{2\pi t}{n}(j+k)i}+e^{\frac{2\pi t}{n}(k-j)i}-e^{\frac{2\pi t}{n}(j-k)i}-e^{\frac{2\pi t}{n}(-j-k)i}
$$
$$
=\frac{1}{4i}\sum_{t=1}^{n}[\cos(\frac{2\pi t}{n}(j+k))+i\sin(\frac{2\pi t}{n}(j+k))+\cos(\frac{2\pi t}{n}(k-j))+i\sin(\frac{2\pi t}{n}(k-j))
$$
$$
-\cos(\frac{2\pi t}{n}(j-k))-i\sin(\frac{2\pi t}{n}(j-k))-\cos(\frac{2\pi t}{n}(-j-k))-i\sin(\frac{2\pi t}{n}(-j-k))]
$$
$$
=\frac{1}{4i}\sum_{t=1}^{n}[\cos(\frac{2\pi t}{n}(j+k))+i\sin(\frac{2\pi t}{n}(j+k))+\cos(\frac{2\pi t}{n}(j-k))-i\sin(\frac{2\pi t}{n}(j-k))
$$
$$
-\cos(\frac{2\pi t}{n}(j-k))-i\sin(\frac{2\pi t}{n}(j-k))-\cos(\frac{2\pi t}{n}(j+k))+i\sin(\frac{2\pi t}{n}(j+k))]
$$
$$
=\frac{1}{2}\sum_{t=1}^{n}\sin(\frac{2\pi t}{n}(j+k))-\sin(\frac{2\pi t}{n}(j-k))
$$
$$
\sum_{t=1}^{n}\sin(\frac{2\pi t}{n}(j+k))-\sin(\frac{2\pi t}{n}(j-k))=0\text{ as }n\text{ is even.}
$$
$$
\sum_{t=1}^{n}\cos (2\pi t \frac{j}{n}) \sin (2\pi t \frac{k}{n})=0 .
$$