---
title: "Introduction to Time Series"
author: Hyoeun Lee
output:  
  pdf_document:
    fig_width: 5
geometry: margin=1in
date: 'Module 1 - Week 1'
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(xts)
library(astsa)

```

# Time Series Elements (Lecture 1.1.1)

### General Learning outcome of this course
 Understand and model the stochastic mechanism that generates an observed time series, and predicts or forecast its future values.

### Time Series Definition

Time Series: a Discrete Stochastic Process

Stochastic Process: collection of Random Variables.
 Denote: $\{{X}_t\}_{t\geq 0}$, $t$: time index
 
* E.g. Discrete-time stochastic Process
	+ Humidity in Illini Union at 3pm on day $t$
	+ Amazon stock price at closing on day $t$
* E.g. Continuous-time stochastic Process
	+ Humidity in Illini Union at time instant $t$
	+ Amazon stock price at time instant $t$


### Time Series Features

* Data collected over time.
* Data cannot be assumed independent. 
* Time series models incorporate **time dependence**

### Many applications

* Economics (monthly unemployment figures, weekly interest rates, daily stock market values); 
* Meteorology (annual precipitation and drought index, global annual temperature series); 
* Agriculture (annual crop production,  food consumption, cultivated area);
* Biological Sciences (cell tissue hourly growth rate, daily seed germination rate); 
* Ecology (annual species abundance, monthly deforestation rate, annual carbon stock rate);
* Medicine (daily blood pressure measurements, brain MRI time series, number of annual breast cancer cases)

### Two approaches to time series analysis:

* Time domain: Modeling a future value of a time series as a parametric function of the current and past values (Box and Jenkins, 1970 and more recently Box, Jenkins and Reinsel, 2013)
* Frequency domain: Assumes the primary feature of a time series is its periodical or systematic sinusoidal variations. These variations are naturally found in most data sets.
* The best path to analyze most data sets is to use both approaches in a complementary fashion.




\newpage
# Time Series Data (Lecture 1.1.2)


### Example 1.1
```{r, example1.1}
par(mfrow=2:1)
tsplot(jj, ylab="QEPS", type="o", col=4, main="Johnson & Johnson Quarterly Earnings")
tsplot(log(jj), ylab="log(QEPS)", type="o", col=4)
# ?jj #R code to load information 
```

*primary patterns?*

* increasing underlying trend and variability
* regular oscillation


\newpage

### Example 1.2
```{r, example1.2}
culer = c(rgb(217,77,30,160,max=255), rgb(30,170,217,160,max=255))
tsplot(gtemp_land, col = culer[1], lwd=2, type="o", pch=20,
       ylab="Temperature Deviations", main="Global Warming")
lines(gtemp_ocean, col=culer[2], lwd=2, type="o", pch=20)
legend("topleft", col=culer, lty=1, lwd=2, pch=20, legend=c("Land Surface", "Sea Surface"), bg="white")
```


*primary patterns?*

* upward trend in both series from 1960-2017 (climate change)
* trend is not linear (sharp upward trend during 1960-2017)



\newpage
### Example 1.3

```{r, example1.3}
djia_return = diff(log(djia$Close))[-1] # R_t = log(X_t) - log(X_{t-1})
par(mfrow=2:1)
plot(djia$Close, col=4)
plot(djia_return, col=4)

```


* $X_t$: DJIA closing value on day $t$
* $R_t$, return, percent of change 

\begin{align}
R_t=(X_t-X_{t-1})/X_{t-1}\\
X_{t-1} (1+ R_t ) = X_t, \quad 1+R_t = X_t / X_{t-1}\\
\end{align}

We approximate $R_T$ with $\log(1+R_t)$, so we say

$$R_t \approx log (X_t) - log (X_{t-1})$$
 
*primary patterns?*

* Around 2008 (financial crisis), return data volatility cluster
* how to forecast volatility of future return: common problem for financial data



 




\newpage
### Example 1.4
```{r, example1.4}
par(mfrow = c(2,1))
tsplot(soi, ylab="", xlab="", main="Southern Oscillation Index", col=4)
text(1970, .91, "COOL", col="cyan4")
text(1970,-.91, "WARM", col="darkmagenta")
tsplot(rec, ylab="", main="Recruitment", col=4) 
```




*primary patterns?*

* annual cycle
* slower frequency of three to seven years (maybe hard to see here) 
* Related to seasonal models





\newpage

# Time Series Models
	
* Why do we model? To reduce uncertainty.
* There is a  probability law that governs each stochastic process.
* A model is an approximation of such a probability law.
* A model is estimated (usually) using past data.

* Three statistics impotant for modeling: Mean, Variance, Dependence




### Model Building Strategy

1. Model Specification: Select the appropriate model for the observed data set
2. Model Fitting: Finding the best estimates of model parameters
3. Model Diagnostics: How well does the model fit the Data?


\newpage
## Some Basic Models


### White Noise Model

* Collection of uncorrelated random variables, $W_t$.
* $W_t$: Random variable with mean 0 and finite variance $\sigma_W^2$.
* Notation: $W_t \sim wn(0,\sigma_W^2)$
* Used as a model for noise in engineering applications

```{r, example1.7}

w = rnorm(500) # 500 N(0,1) variates
tsplot(w, col=4, main="white noise")
```


\newpage
### Moving Average (MA) Model

* Moving average of white noise
* $X_t = \frac{1}{2} (W_{t-1} + W_t )$
* Interpret the time series as linear combination of white noise
* could be used for smoothing (in Section 3.3)

```{r, example1.8}
v = filter(w, sides=1, filter=rep(1/2,2)) # moving average
tsplot(v, ylim=c(-3,3), col=4, main="moving average")
```

\newpage
### Autoregression (AR) Model

* $X_t = 1.5 X_{t-1} -0.75 X_{t-2} + W_t$
* Regression or prediction of the current value ($X_t$) as a function of the past two values.
* periodic behavior


```{r, example1.9}
set.seed(90210)
w = rnorm(250 + 50) # 50 extra to avoid startup problems
x = filter(w, filter=c(1.5,-.75), method="recursive")[-(1:50)]
tsplot(x, main="autoregression", col=4)
```

\newpage
### Random Walk with drift

* $X_t = \delta + X_{t-1} + W_t$
* Random walk: value at time $t$ is the value of the series at time $t-1$ plus a completely random movement determined by $W_t$. 
* $\delta$: drift
* With initial condition $x_0=0$, $$X_t=\delta t + \sum_{j=1}^{t} W_j$$

```{r, example1.10}
set.seed(314159265) # so you can reproduce the results
w  = rnorm(200) 
x  = cumsum(w)
wd = w +.3 
xd = cumsum(wd)
tsplot(xd, ylim=c(-2,80), main="random walk", ylab="", col=4)
abline(a=0, b=.3, lty=2, col=4) # drift
lines(x, col="darkred")
abline(h=0, col="darkred", lty=2)

```

\newpage
### Signal Plus noise

* $X_t = 2 \cos (2\pi \frac{t+15}{50}) + W_t$
* $X_t= A \cos (2\pi \omega t+\phi)+W_t$
* $A$: amplitude, $\omega$: frequency, $\phi$: phase shift. 


```{r, example1.11}
cs = 2*cos(2*pi*(1:500)/50 + .6*pi)
w  = rnorm(500,0,1) 
par(mfrow=c(3,1), mar=c(3,2,2,1), cex.main=1.5)   # help(par) for info
tsplot(cs, col=4, ylab="", main = expression(x[t]==2*cos(2*pi*t/50+.6*pi)))
tsplot(cs + w, col=4,ylab="", main = expression(x[t]==2*cos(2*pi*t/50+.6*pi)+N(0,1)))
tsplot(cs + 5*w, col=4,ylab="", main = expression(x[t]==2*cos(2*pi*t/50+.6*pi)+N(0,25)))
```



